R for Clinical Study Reports and Submission

### 2021-08-29 tlf-baseline.Rmd
update content structure
update tlf-baseline table

## Table for Baseline Characteristics

```{r}
library(table1)
library(r2rtf)
data(r2rtf_adsl)

tbl <- table1(~ SEX + AGE + RACE | TRT01P, data = r2rtf_adsl)
tbl
```

```{r}
tbl <- tbl %>% as.data.frame() %>% 
               as_tibble() %>% 
               mutate( across(everything(), 
                        ~ stringr::str_replace_all(.x, intToUtf8(160), " ") ) 
                      )
```

```{r}
tbl %>% 
      rtf_body() %>% 
      rtf_encode() %>% 
      write_rtf("tlf/bs-1.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/bs-1.pdf")
```

>>>>

# Table for Baseline Characteristics

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize critical demographic and baseline characteristics of the patients
in the section 11.2, demographic and other baseline characteristics of a CSR.

In this chapter, we illustrate how to create a simplified 
baseline characteristics table in a study.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_base.pdf")
```

There is many R packages that can effciently summarize baseline information.
The [`table1`](https://github.com/benjaminrich/table1) R package is one of them.

```{r}
library(table1)
library(r2rtf)
library(haven)
library(stringr)
```

As in previous chapters, we first read `adsl` dataset that contain all 
required information for baseline characteristics table.

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
```

For simplicity, we only analyze `SEX`, `AGE` and `RACE` in this example
using the `table1` R package. 
More details of the `table` R package can be found in the package
[vignettes](https://benjaminrich.github.io/table1/vignettes/table1-examples.html)

The `table1` R package directly create an HTML report.

```{r}
tbl <- table1(~ SEX + AGE + RACE | TRT01P, data = adsl)
tbl
```

The code below transfer the output into a dataframe 
that only contain ASCII character 
recommended by regulatory agencies. 
`tbl_base` is used as input for `r2rtf` to create final report.

```{r}
tbl_base <- tbl %>% 
              as.data.frame() %>% 
              mutate( across(everything(), 
                      ~ str_replace_all(.x, intToUtf8(160), " ") ) 
                    )

names(tbl_base) <- str_replace_all(names(tbl_base), intToUtf8(160), " ")
tbl_base
```

We start to define the format of the output. We highlighted items that 
is not discussed in previous discussion.

`text_indent_first` and `test_indent_left` are used to control the indent
space of text. They are helpful when you need to control the white space 
of a long sentence. For example, "AMERICAN INDIAN OR ALASKA NATIVE" 
in the table.

```{r}
colheader1 <- paste(names(tbl_base), collapse = "|")
colheader2 <- paste(tbl_base[1, ], collapse = "|")
rel_width <- c(2.5, rep(1, 4))

tbl_base[-1, ] %>% 
      rtf_title("Participant Baseline Characteristics", 
                "(All Participants Randomized)") %>%
      rtf_colheader(colheader1, 
                    col_rel_width = rel_width) %>% 
      rtf_colheader(colheader2, 
                    border_top = "",
                    col_rel_width = rel_width) %>%
      rtf_body(col_rel_width = rel_width,
               text_justification = c("l", rep("c", 4)), 
               text_indent_first = -240,
               text_indent_left = 180) %>% 
      rtf_encode() %>% 
      write_rtf("tlf/tlf_base.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_base.pdf")
```
